---
title: "SDG Combining report"
author: "Michael Rahija"
date: "September 6, 2019"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# Report to combine datasets


## Set up
```{r, warning = F, message = F }
#set up
knitr::opts_chunk$set(echo = T, warning=F, message = F)
library(FAOSDGdata)
library(dplyr)

```

## Download single or multiple datasets and load as list(s)

To get a singe datasets you can use the getData function. 
```{r }
sdg <- getData("2.1.1")

head(sdg)
```

To retrieve multiple datasets, you can use a vector of the corresponding SDG dataset numbers, 
or to get all, you can set all = T.
```{r, cache = T, wannings = F}
allSdgs <- getDatasets(all = T)

someSDGs <- getDatasets(c("2.1.1", "2.1.2"))

head(someSDGs)
```


## Merging multiple datasets into single dataframe

Different technical units within FAO provide the SDG datasets, and the column names don't always match. Also, there are additional columns for some SDG indicators which are not applicable to others. Accordingly, cleanData subsets the datasets for only the columns which are present across and all SDG datasets. 

```{r }
cleanall <- lapply(allSdgs, cleanData)

Sdgsdf <- do.call("rbind", cleanall)

nrow(Sdgsdf)
```


## Harmonizing SDG dataset(s) with M49

The UNSD M49 standard is the geographic naming standard adopted by FAO for SDG data dissemination. However, this standard changes frequently, so scrape_m49 was written to scrape in real time the M49 standard from the UNSD website.

**FAO has some slight deviations to the M49 standard which are applied with the function adopt_m49_FAO().**

Also, to use the M49 to harmonize the names in the SDG dataset, it must be flattened so it can be merged. 

```{r }
##--Harvest M49, and flatten for merger
geo <- scrape_m49() %>%
          adopt_m49_for_FAO()

flatm49 <- flatten_m49(geo)

```

Sometimes technical units within FAO submit data using geographic areas which may be relevant for their type of analysis, but not align directly with the FAO. Accordingly, those records are removed.  Below is a dataframe showing the records that were removed. 

```{r }

#filter sdg dataset for only areas with valid m49 codes
sdgs_validcodes <- filter(Sdgsdf, areacode %in% flatm49$code)
sdgs_INvalidcodes <-  filter(Sdgsdf, !(areacode %in% flatm49$code))

invalid <- sdgs_INvalidcodes %>%
                      group_by(areacode,areaname) %>%
                      summarize(N = n())

as.data.frame(invalid)
```

After harmonizing names, the table below shows the total records for each geographic area. 

```{r }
#merge sdg dataset with m49 to create consistent area names
flatm49 <- select(flatm49, -level)

x <- merge(sdgs_validcodes,
           flatm49,
           by.x = "areacode",
           by.y = "code",
           all.x = T)

#update areaname variable an drop
x$areaname <- x$name

x <- select(x, -name)

geotable <- x %>%
              group_by(areaname) %>%
              summarize(N = n())
#check names
knitr::kable(geotable)
```

## Harmonizing units of measure
The table below shows the units of measure used across all SDG datasets
```{r }
dirtyunits <- x %>% 
                group_by(unit) %>%
                summarize(N = n())

knitr::kable(dirtyunits, caption = "Before harmonizing")
````

There are some NAs for units which correspond to NA values. They would be excluded when analyzed.
```{r }
x$unit <- harmonizeUnits(x$unit)
cleanunits <- x %>% 
                group_by(unit) %>%
                summarize(N = n())

knitr::kable(cleanunits, caption = "After harmonizing")
````


## Harmonizing/checking indicator names


```{r }
indicators <- x %>% 
                group_by(indicator, indicatorDescription) %>%
                summarize(N = n()) %>%
                arrange(indicator)

knitr::kable(indicators, caption = "Before harmonizing")
````

## Harmonizing/checking indicator names


```{r }
indicators <- x %>% 
                harmonizeIndicators() %>%
                group_by(indicator, indicatorDescription) %>%
                summarize(N = n()) %>%
                arrange(indicator)

knitr::kable(indicators, caption = "Before harmonizing")

x <- harmonizeIndicators(x)
write.csv(x, file = "temp_mergedSDG.csv", row.names = F)
````

## Reviewing how much data for each indicator


```{r }
null.values = c("NA", "N", "Not validated")
y <- x %>% 
  filter(areatype == "3.0-Country") %>%
  group_by(indicator, indicatorDescription) %>%
  mutate(Valid = !(value %in% null.values)) %>%
  filter(Valid == TRUE) %>%
  summarize(Countries = length(unique(areaname)),
            Valid = n(),
            Years = length(unique(timeperiod)))

y <- arrange(y, indicator)
write.csv(y, file = "temp_summary_available_data.csv", row.names = F)
knitr::kable(y)              
````